\LARGE\textbf{Designing Studies and Evaluating Research Results:\\ Type M and Type S Errors for\\ Pearson Correlation Coefficient }

### 2020-06-12 Documents/Paper_main/Paper_main.Rnw
Corrige Giulia 2

We can implement the prospective design analysis using the function \texttt{pro\_r()} which inputs and outputs are displayed in Figure~\ref{fig:pro_r}. About 125 participants are necessary to have 80\% probability to detect an effect of at least $\rho=\pm .25$, if it actually exists. With this sample size, the Type S error is minimized and approximates zero. In this study design, the Type M error is 1.12 indicating that statistically significant results are on average exaggerated by 12\%. It is possible to notice that the critical values are $r=\pm .18$, further highlighting that our plausible effect size is actually included among those values that lead to the acceptance of the alternative hypothesis.

@@

In a design analysis, it is advisable to investigate how the inferential risks would change according to different scenarios in terms of statistical power and plausible effect size. Changes in both these factors impact Type M and Type S errors. For example, maintaining the plausible correlation of $\rho = .25$, if we decrease statistical power from .80 to .60 only 76 participants are required (see Table~\ref{tab:Table_pro_r}). However, this is associated with an increased Type M error rate from 1.12 to 1.28. That is to say, with 76 subjects the plausible effect size will be on average overestimated by 28\%. Alternatively, imagine that we would like to maintain a statistical power of 80\%, what happens if the plausible effect size is slightly larger or smaller? The necessary sample size would spike to 344 for a $\rho=.15$  and decrease to 60 for $\rho=.35$. In both scenarios, the Type M error remains about 1.12, which reflects the more general point that for 80\% power, Type M error is around 1.10. In all these scenario, Type S error is close to zero, hence not worrisome.

>>>>

We can implement the prospective design analysis using the function \texttt{pro\_r()} which inputs and outputs are displayed in Figure~\ref{fig:pro_r}. About 125 participants are necessary to have 80\% probability to detect an effect of at least $\rho=\pm .25$, if it actually exists. With this sample size, the Type S error is minimized and approximates zero. In this study design, the Type M error is 1.11 indicating that statistically significant results are on average exaggerated by 11\%. It is possible to notice that the critical values are $r=\pm .18$, further highlighting that our plausible effect size is actually included among those values that lead to the acceptance of the alternative hypothesis.

@@

In a design analysis, it is advisable to investigate how the inferential risks would change according to different scenarios in terms of statistical power and plausible effect size. Changes in both these factors impact Type M and Type S errors. For example, maintaining the plausible correlation of $\rho = .25$, if we decrease statistical power from .80 to .60 only 76 participants are required (see Table~\ref{tab:Table_pro_r}). However, this is associated with an increased Type M error rate from 1.11 to 1.28. That is to say, with 76 subjects the plausible effect size will be on average overestimated by 28\%. Alternatively, imagine that we would like to maintain a statistical power of 80\%, what happens if the plausible effect size is slightly larger or smaller? The necessary sample size would spike to 344 for a $\rho=.15$  and decrease to 60 for $\rho=.35$. In both scenarios, the Type M error remains about 1.12, which reflects the more general point that for 80\% power, Type M error is around 1.10. In all these scenario, Type S error is close to zero, hence not worrisome.
